Backward designing a reproducible framework for undergraduate data science research

### 2019-12-05 Rmd/discussion.Rmd
fixed methods & recompiled pdf (with line numbers)

1. build self-confidence, proficiency in data analysis and statistics

1. greater interest in quantitative careers and/or grad school in quant fields

1. Working with real data to solve real scientific questions

1. Learn by example interplay between data science (statistics) and science research

Benefits of our framework include enhanced

## How does the framework relate to ideas in Nolan and Temple Lang 2010??

@@

Our measures of students' self-confidence in research ability was merely subjective. In 
future iterations of our framework, we would like to measure systematic and objective 
outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, 
both before and after the mentored research project. We would use validated survey 
questions that focused on student beliefs about themselves, their skills, and their 
future careers. **maybe list resources for such survey questions**

## Areas for improvement

1. reproducible research best practices

@@

1. survey questions might include those from @vance2017eight. I have the survey as a pdf in mendeley. It is part of the supp info for @vance2017eight.

>>>>

The student test cases for our framework demonstrated greater 
self-confidence and greater proficiency in data science skills over the 
course of the research projects. They used real-world data sources to 
address real scientific research questions.
Additionally, they showed great interest
in quantitative and data science careers. After graduation, one student immediately enrolled
in statistics graduate training, while the other sought employment in 
health care analytics.

@@

From our current perspective, we offer a number of framework critiques and opportunities for improvement. Our measure of students' self-confidence in research ability was merely subjective. In 
future iterations of our framework, we would like to measure systematic and objective 
outcomes. One strategy for implementing this is to administer a survey, including questions from @vance2017eight, 
both before and after the mentored research project. We would use validated survey 
questions that focused on student beliefs about themselves, their skills, and their 
future careers.

@@

### Framework development with backward design

In future research, we will continue to develop our framework for 
undergraduate data science research by explicitly 
incorporating backward design principles [@wiggins2005understanding]. 
Following @wiggins2005understanding, we will identify desired results, 
determined acceptable evidence, and planned learning experiences.

Before identifying desired results, we will prioritize topics from 
@nolan2010computing. Specifically, we will assign all terms from Figure 1 
of @nolan2010computing into one of three categories:

1. worth being familiar with 
1. important to know and do
1. enduring understanding

We've tabulated below the @nolan2010computing terms for the current 
framework and its student projects.

```{r table1, echo = FALSE, message = FALSE}
library(magrittr)
circles <- tibble::tibble(Class = 1:3, Circle = c("Worth being familiar with", "Important to know and do", "Enduring understanding"))
readr::read_csv("../data/nolan-figure1-circles.csv") %>%
  dplyr::left_join(circles) %>%
  dplyr::arrange(desc(Class)) %>%
  dplyr::select(- Class) %>%
  gt::gt() %>%
  gt::tab_header(title = "Prioritizing Key Terms from Figure 1 of @nolan2010computing", 
                 subtitle ="xxx" ) %>%
  gt::as_latex()
```

Potential benefits of incorporating backward design ideas include clearer 
articulation of goals and better assessment of goal achievement.

We see our framework as one contribution to scholarship on improving data 
science training programs. Given the increasing economic need, in the USA 
and abroad, for 
data scientists and other researchers with quantitative training, we 
anticipate that our framework and its future iterations will continue
to prepare students for data science careers by offering training in tangible and 
transferable analytic skills in the context of solving scientific questions.

### 2019-12-05 Rmd/jse-2019-ms.Rmd
fixed methods & recompiled pdf (with line numbers)

---
title: 'Backward designing a reproducible framework for undergraduate data science research'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
output: 
  - bookdown::pdf_document2
  - bookdown::html_document2
---

>>>>

---
title: 'Backward designing a reproducible framework for undergraduate data science research'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
output: 
  bookdown::pdf_document2:
    includes:
      in_header: head.tex
---

### 2019-12-05 Rmd/methods.Rmd
fixed methods & recompiled pdf (with line numbers)

We used backward design principles to create and implement a 
framework for mentored undergraduate data science research 
projects with big data. Below, we describe our use of 
backward design and ideas from @nolan2010computing.
We then detail our framework and its initial implementation.

## Framework design

We started with the backward design principles from @wiggins2005understanding. We prioritized

@@

**justify each step**

**Maybe use/cite Box's science and statistics article from... 1976?**

1. Overview of our framework
  a. student-led, hypothesis-driven research
  b. BH and FB helped with data acquisition
  c. helped with research question formulation
  d. how to describe roles for students v mentors?

@@

1. State the questions and research hypotheses that the two students chose

1. Data description

- how did we/they conceive of projects?

- how did we balance our input with that of the students?
- What data did we use?
- How did we collect data? (which packages, code to sample twitter). How did Jinyu get her stock market data? 
- What is the sampling scheme - for the 1% sample - from the entirety of Twitter? (Cite the webpage that documents the API)

>>>>

We designed and implemented a 
framework for mentored undergraduate data science research 
projects with big data. Below, we describe our initial framework 
and connect it to ideas from @nolan2010computing.

@@

@@
